---
title: "Global_Exploratory"
output: html_document
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, message=FALSE)
```

```{r}
#user info
if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis/"
if(Sys.info()[["user"]] == "robmarty") dropbox_file_path <- "~/Dropbox/World Bank/Side Work/COVID Social Media Analysis"


#packages
library(tidyverse)

data <- read_rds(file.path(dropbox_file_path, "Data/google_trends/FinalData/gtrends_full_timeseries/gtrends_otherdata_varclean.rds"))

data_cor_02 <- read_rds(file.path(dropbox_file_path, "Data/google_trends/FinalData/gtrends_full_timeseries/correlation_datasets/correlations_gtrends_otherdata_varclean_since2020-02-01.rds"))

data_wdi'

```


# Ungroup variables (which are grouped by geo)
```{r}
data_cor_02 <- 
  data_cor_02 %>% 
  ungroup(geo) 

data_cor_02 %>% count(keyword_en, sort = TRUE)
```

# We create categories for key words

```{r}
data_cor_02 <- 
  data_cor_02 %>% 
  mutate(
    categories = case_when(
      keyword_en %in% 
        c("divorce", "debt",  "unemployment", "telework",
          "social isolation", "unemployment insurance", "unemployment benefits", 
          "unemployment office", "file for unemployment", "loan",
          "online education") ~ "potential_consequences",
      keyword_en %in% 
        c("therapy","anxiety", "stress", "panic", "sadness", "fatigue", "insomnia", 
          "suicide", "depression", "loneliness", "worry", "boredom") ~ "mental_health",
      keyword_en %in% 
        c("chloroquine", "doctor", "ventilators", "face mask", "coronavirus test", 
          "thermometers", "doctors", "hospital", "hydroxychloroquine") ~ "treatment",
      keyword_en %in% 
        c( "fever", "cough", "pneumonia", "coronavirus symptoms", "tired", "cough",
           "sintomas do coronavirus", "shortness of breath", "covid symptoms", 
           "sore throat", "anosmia", "corona symptoms", "loss of smell",
           "loss of taste", "what are the symptoms of coronavirus", "ageusia", 
           "i have coronavirus", "coronavirus help", "i have a fever", "i have a cough",
           "fever treatment", "my eyes hurt", "how to treat coronavirus", "i can't smell", 
           "i can't taste", "difficulty breathing", "breathing") ~ "symptoms",
      keyword_en %in% 
        c("wash hands","stay at home", "social distance", "i stay at home",
          "disinfectants") ~ "prevention", 
      keyword_en%in% c("coronavirus", "covid-19") ~ "coronavirus_general", 
      keyword_en %in% 
        c("vaccine", "covid vaccine", "covid vaccine side effects",
          "covid vaccine dangerous", "covid vaccine safety", "safety of covid vaccine", 
          "covid vaccine change dna", "does covid vaccine change dna", 
          "long term effects of covid vaccine", "can covid vaccine cause infertility", 
          "is the covid vaccine the mark of the beast", "covid vaccine cause infertility", 
          "does covid vaccine cause infertility", "is covid vaccine the mark of the beast", 
          "negative side effects of covid vaccine") ~ "vaccine"
      
      )
  )
```


#Overall Correlation & Correlation Lag 

For this section, we want to keep words that appear at least 20 times, to prevent correlations based on very few observations; 
these excluded words refer mostly to vaccine-related searches, which are not relevant for the prediction efforts. 
```{r}
common_keywords <- 
  data_cor_02 %>% 
  count(keyword_en, sort = TRUE) %>% 
  filter(n > 20, !is.na(keyword_en)) %>% 
  pull(keyword_en)

data_cor_02 <- 
  data_cor_02 %>% 
  filter(keyword_en %in% common_keywords)
```


## Overall correlation of terms & associated lag -- Cases
```{r}
data_cor_02 %>% 
  filter(type == "Cases") %>% 
  group_by(keyword_en) %>% 
  summarize(
    mean_cor = mean(cor, na.rm = TRUE),
    mean_lag = mean(lag, na.rm = TRUE)
  ) %>% 
  arrange(desc(mean_cor))
```

## Overall correlation of terms & associated lag -- Deaths

```{r}
data_cor_02 %>% 
  filter(type == "Deaths") %>% 
  group_by(keyword_en) %>% 
  summarize(
    mean_cor = mean(cor, na.rm = TRUE),
    mean_lag = mean(lag, na.rm = TRUE)
  ) %>% 
  arrange(desc(mean_cor))
```

## Overall correlation & lag by category

```{r}
data_cor_02 %>% 
  filter(type == "Deaths") %>% 
  group_by(categories) %>% 
  summarize(
    mean_cor = mean(cor, na.rm = TRUE),
    mean_lag = mean(lag, na.rm = TRUE)
  ) %>% 
  arrange(desc(mean_cor))
```


#Underreporting (Highlighting the cases we mentioned + others)



#Look at other variables that might explain where correlation is high v. not so high

```{r}
data_cor_02 %>% count(geo, 
data %>% count(geo, income, 
```

