---
title: "01_Cleaning"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggrepel)
library(ggcorrplot)
library(corrplot)
library(Hmisc)
library(PerformanceAnalytics)
library(stringr)

#user file
if(Sys.info()[["user"]] == "wb537287") user <- "/Users/wb537287/"

#folder path
data_folder <- paste0(user, "Dropbox/COVID Social Media Analysis/Data/usa_vaccine/RawData/")
elec_folder <- paste0(user, "Dropbox/COVID Social Media Analysis/Data/usa_elections/")
gtrends_folder <- paste0(user, "Dropbox/COVID Social Media Analysis/Data/google_trends/FinalData/gtrends_regional/")
output_folder <- paste0(user, "Dropbox/COVID Social Media Analysis/Paper Figures/Vaccine_Analysis/")

#file_path
file_doses_adm <- paste0(data_folder, "us-daily-covid-vaccine-doses-administered.csv")
file_doses_dis <- paste0(data_folder, "us-total-covid-vaccine-doses-distributed.csv")
file_share_used <- paste0(data_folder, "us-share-covid-19-vaccine-doses-used.csv")
file_share_vac <- paste0(data_folder, "us-covid-19-share-people-vaccinated.csv")
file_gtrends <- paste0(gtrends_folder, "gtrends_regional.Rds")
file_elec <- paste0(elec_folder, "data_elec_cleaned.csv")


data_doses_adm <- 
  file_doses_adm %>% 
  read_csv()

data_doses_dis <- 
  file_doses_dis %>% 
  read_csv()

data_share_used <- 
  file_share_used %>% 
  read_csv()

data_share_vac <- 
  file_share_vac %>% 
  read_csv()

data_gtrends <- 
  file_gtrends %>% 
  read_rds()

data_elec <- 
  file_elec %>% 
  read_csv()

#the data of share of people vaccinated as of June 18
data_share_vac_jun18 <- 
  data_share_vac %>% 
  filter(Day == "2021-06-18")

#removing duplicates from gtrends
data_gtrends <- 
  data_gtrends %>% 
  group_by_all() %>%
  filter(n() == 1)

```

# Merging the 3 vaccine-related datasets

```{r}
data_vaccine <- 
  data_doses_adm %>% 
  left_join(data_doses_dis, by = c("Entity", "Day", "Code")) %>% 
  left_join(data_share_used, by = c("Entity", "Day", "Code")) %>% 
  left_join(data_share_vac, by = c("Entity", "Day", "Code"))
```

#Merging vaccine and search data

vaccine data is daily; search data is in 2 periods: 7 months (Dec,20-May 21) and 3 months (March-May 21)

To merge, we can: 
a) Create a variable of average/median doses for those 2 periods, then calculate correlations

```{r}
data_vac_mean <- 
  data_vaccine %>% 
  group_by(Entity) %>% 
  summarise(
    sum_vaccinations = sum(daily_vaccinations, na.rm = TRUE),
    total_distributions = max(total_distributed, na.rm = TRUE), 
    across(daily_vaccinations:share_doses_used, ~mean(., na.rm = TRUE))
  )

```

Merge vaccine and search data
```{r}
data <- 
  data_gtrends %>% 
  left_join(data_vac_mean, by = c("location" = "Entity"))
```

Merging google trends data with the share of people vaccinated as of June 18

```{r}
data_jun18 <- 
  data_gtrends %>% 
  left_join(data_share_vac_jun18, by = c("location" = "Entity"))
```

We create an indicator variable for whether there is a search or a missing value 
```{r}
data_jun18 <- 
  data_jun18 %>% 
  mutate(
    missing_search = is.na(hits), 
    search = if_else(missing_search == FALSE, "yes", "no")
  )
```


# Analysis

we now need to decide what works to look at in order to evaluate their correlation with share of doses used
We start with a few keywords just to check

```{r}
data %>% 
  ungroup %>% 
  count(keyword)
```

We select the keywords related to the vaccine or potential correlations with the anti-vaccine movement
```{r}
keywords_vaccine <- 
c(
  "can covid vaccine cause infertility",
  "biden election fraud",
  "biden stole the election",
  "bill gates microchip",
  "covid microchip",	
  "covid symptoms",				
  "covid vaccine",				
#  "covid vaccine austism",			
#  "covid vaccine cause autism",				
  "covid vaccine cause infertility",				
  "covid vaccine change dna",				
  "covid vaccine dangerous",				
  "covid vaccine infertility",				
  "covid vaccine magnetic",				
#  "covid vaccine mark of the best",				
  "covid vaccine microchip",				
  "covid vaccine safety",			
  "covid vaccine side effects",	
  "does covid vaccine cause infertility",				
  "does covid vaccine change dna",
  "election fraud",	
  "election stolen",	
  "fox news",	
  "hunter biden laptop",	
  "is covid vaccine the mark of the beast",			
  "is the covid vaccine the mark of the beast",	
  "long term effects of covid vaccine",	
  "negative side effects of covid vaccine",	
#  "newsmax",	
  "parler",
  "qanon",
  "safety of covid vaccine",	
  "stop the count",	
  "vaccine",				
  "vaccine cause autism"
  )
```


```{r}
# we filter for the broader time span; we convert NAs to 0 
data %>% 
  filter(time_span == "2020-12-01_2021-05-31", keyword %in% keywords_vaccine) %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  ggplot(aes(share_doses_used, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0,100))
```

Change the time span to only focus on March-May

```{r}
data %>% 
  filter(time_span == "2021-03-01_2021-05-31", keyword %in% keywords_vaccine) %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  ggplot(aes(share_doses_used, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0,100))
```

- we find that there isn't much of a change from switching the time span from dec-march to may-march

# Just focusing on "Covid vaccine side effects" 

We look at the states with covid vaccine side effects
```{r}
data %>% 
  filter(time_span == "2020-12-01_2021-05-31", keyword == "covid vaccine side effects") %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  ggplot(aes(share_doses_used, hits)) + 
  geom_point() +
  geom_smooth() + 
  geom_text_repel(aes(label = location)) + 
  coord_cartesian(ylim = c(0,100))
```


# Changing dependent variable to share of people vaccinates

```{r}
# we filter for the broader time span; we convert NAs to 0 
data_jun18 %>% 
  filter(time_span == "2020-12-01_2021-05-31", keyword %in% keywords_vaccine) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0,100))
```


```{r}
data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% 
      c(
        "covid vaccine",
        "covid vaccine safety",
        "covid vaccine side effects", 
        "covid vaccine cause infertility",
        "fox news",
        "is the covid vaccine the mark of the beast",
        "long term effects of covid vaccine",
        "newsmax"
      )
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0, 100))

```

```{r}
data_jun18 %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% 
      c(
        "covid vaccine",
        "covid vaccine safety",
        "covid vaccine side effects", 
        "covid vaccine cause infertility",
        "fox news",
        "is the covid vaccine the mark of the beast",
        "long term effects of covid vaccine",
        "newsmax",
        "vaccine"
      )
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0, 100))
```


#Histogram of mark of the beast

```{r}
mark_beast <- 
  data_jun18 %>% 
  filter(keyword == "is the covid vaccine the mark of the beast") %>% 
  ggplot(aes(people_vaccinated_per_hundred, group = search, color = search, fill = search)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Searches for 'Is covid-19 the mark of the beast' across US states", 
    subtitle = "States with searches for fake news have lower proportion of vaccinated people", 
    x = "Percentage of people vaccinated"
  )

ggsave(mark_beast, filename = paste0(output_folder, "mark_beast.png"))

```


```{r}
data_jun18 %>% 
  filter(keyword == "long term effects of covid vaccine") %>% 
  ggplot(aes(people_vaccinated_per_hundred, group = search, color = search, fill = search)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Searches for 'long term effects of covid vaccine' across US states"
  )
```

#Focus on vaccine-related searches - for % of vaccinated people

```{r}
data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) 
```

```{r}
vaccine_related <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("vaccine", "covid vaccine side effects","covid vaccine safety")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states have more vaccine-related searches", 
    subtitle = "Searches for vaccine safety or side effects are higher in states with greater share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  )

ggsave(vaccine_related, filename = paste0(output_folder, "vaccine_related.png"), width = 8)

```


## Limiting to safety and side effects 

```{r}
vaccine_safety <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects","covid vaccine safety")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more about the safety of the vaccine", 
    subtitle = "Searches for vaccine safety or side effects are higher in states with greater share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  )

ggsave(vaccine_safety, filename = paste0(output_folder, "vaccine_safety.png"), width = 8)

```

#Focus on anti-vaxx searches

```{r}
antivax <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% 
      c(
        "fox news",
        "is the covid vaccine the mark of the beast",
        "newsmax"
    )
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "Searches across US states", 
    subtitle = "Searches for anti-vaxx or fox news are negatively correlated with share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  )

ggsave(antivax, filename = paste0(output_folder, "antivax.png"), width = 10)

```

#Only vaccine searches

```{r}
vaccine <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("vaccine")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  geom_text_repel(aes(label = location)) + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more for the vaccine", 
    subtitle = "There is a clear positive correlation between vaccine searches and share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  )

ggsave(vaccine, filename = paste0(output_folder, "vaccine.png"), width = 8)

```


# All keywords
```{r}
all_terms <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% keywords_vaccine
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) +
  labs(
    title = "Vaccine-related searches and share of vaccinations across US states", 
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  ) + 
  coord_cartesian(ylim = c(0, 100))

ggsave(all_terms, filename = paste0(output_folder, "all_terms.png"), height = 9, width = 10)

```

# Vaccine Analysis

## Merging election and vaccine data

```{r}
#clean data_elec
data_elec <- 
  data_elec %>% 
  filter(winner == 1) %>% 
  mutate(state = str_to_title(state)) %>% 
  mutate(state = if_else(state == "District Of Columbia", "District of Columbia", state))


#merge data_jun18 and data_elec
data_jun18 <- 
  data_jun18 %>% 
  left_join(data_elec, by = c("location" = "state"))
```

## We compute the R^2

```{r}
data_jun18_vaccine <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("vaccine")
  )

data_jun18_side_effects <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects")
  )

data_jun18_safety <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine safety")
  )

rsq <- function(x, y) summary(lm(y~x))$r.squared

rsq_vaccine = 
  rsq(data_jun18_vaccine$people_vaccinated_per_hundred, data_jun18_vaccine$hits) %>% 
  round(digits = 2)

rsq_side_effects = 
  rsq(data_jun18_side_effects$people_vaccinated_per_hundred, data_jun18_side_effects$hits) %>% 
  round(digits = 2)


rsq_safety = 
  rsq(data_jun18_safety$people_vaccinated_per_hundred, data_jun18_safety$hits) %>% 
  round(digits = 2)
```

## Graphs

```{r}
vaccine_elections <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("vaccine")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point(aes(color = as.factor(party_simplified)), size = 2) +
  geom_smooth(method = "lm") + 
  geom_text_repel(aes(label = location)) + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more for the vaccine", 
    subtitle = "There is a clear positive correlation between vaccine searches and share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches", 
    color = "Winner in 2020 Presidential Elections"
  )

ggsave(vaccine_elections, filename = paste0(output_folder, "vaccine_elections.png"), height = 9, width = 10)
```

```{r}
vaccine_elections_rsq <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("vaccine")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point(aes(color = as.factor(party_simplified)), size = 2) +
  geom_smooth(method = "lm") + 
  geom_text_repel(aes(label = location)) + 
  annotate(geom="label", x=40, y=90, label = paste0("R^2 = ", rsq_vaccine), color="brown4", size = 8) +
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more for the vaccine", 
    subtitle = "There is a clear positive correlation between vaccine searches and share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches", 
    color = "Winner in 2020 Presidential Elections"
  )

ggsave(vaccine_elections_rsq, filename = paste0(output_folder, "vaccine_elections_rsq.png"), height = 9, width = 10)
```

```{r}
vaccine_elections_safety <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects","covid vaccine safety")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point(aes(color = as.factor(party_simplified)), size = 2) +
  geom_smooth(method = "lm") + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more for the vaccine", 
    subtitle = "There is a clear positive correlation between vaccine searches and share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches", 
    color = "Winner in 2020 Presidential Elections"
  )

ggsave(vaccine_elections_safety, filename = paste0(output_folder, "vaccine_elections_safety.png"), height = 9, width = 10)
```


```{r}
vaccine_safety <- 
  data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects","covid vaccine safety")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more about the safety of the vaccine", 
    subtitle = "Searches for vaccine safety or side effects are higher in states with greater share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches"
  )

ggsave(vaccine_safety, filename = paste0(output_folder, "vaccine_safety.png"), width = 8)

```

```{r}
rsq_text <- data.frame(
  label = c(paste0("R^2 = ", rsq_safety),paste0("R^2 = ", rsq_side_effects)),
  keyword   = c("covid vaccine safety", "covid vaccine side effects")
)

vaccine_safety_rsq <- 
data_jun18 %>% 
  filter(
    time_span == "2020-12-01_2021-05-31",
    keyword %in% c("covid vaccine side effects","covid vaccine safety")
  ) %>% 
  ggplot(aes(people_vaccinated_per_hundred, hits)) + 
  geom_point(aes(color = as.factor(party_simplified)), size = 2) +
  geom_smooth(method = "lm") + 
  geom_label(
    data = rsq_text,
    mapping = aes(x = c(45, 45), y = c(82, 90), label = label),
    size = 6, 
    color="brown4"
  ) +
  facet_wrap(vars(keyword), scales = "free") +
  labs(
    title = "More vaccinated states search more for the vaccine", 
    subtitle = "There is a clear positive correlation between vaccine searches and share of vaccinations",
    x = "Percentage of people vaccinated for each state", 
    y = "Hits in Google Searches", 
    color = "Winner in 2020 Presidential Elections"
  )

ggsave(vaccine_safety_rsq, filename = paste0(output_folder, "vaccine_safety_rsq.png"), height = 9, width = 10)
```

## Correlations


# APPENDIX

# Changing dependent variable to sum of daily vaccinations/maximum distributed

We'd like to do sum of daily vaccinations divided by maximum of distributed
```{r}
data <- 
  data %>% 
  mutate(vaccination_usage = sum_vaccinations/total_distributions)
  

# we filter for the broader time span; we convert NAs to 0 
data %>% 
  filter(time_span == "2020-12-01_2021-05-31") %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  ggplot(aes(vaccination_usage, hits)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0,100))
```


```{r}
data %>% 
  filter(time_span == "2020-12-01_2021-05-31", keyword == "covid vaccine side effects") %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 
  ggplot(aes(vaccination_usage, hits)) + 
  geom_point() +
  geom_smooth() + 
  geom_text_repel(aes(label = location)) + 
  coord_cartesian(ylim = c(0,100))
```


#Correlation Plot of Keywords


```{r}

data_jun18_wide <- 
  data_jun18 %>% 
  filter(time_span == "2020-12-01_2021-05-31") %>% 
  mutate(
    hits = replace(hits, is.na(hits), 0), 
    keyword = word(keyword,-2)
  ) %>% 
  group_by(location) %>% 
  summarise(keyword, hits) %>% 
  ungroup() %>% 
  pivot_wider(names_from = keyword, values_from = hits)

data_jun18_wide <- as.data.frame(data_jun18_wide)

row.names(data_jun18_wide) <- data_jun18_wide$location

data_jun18_wide <- data_jun18_wide %>% select(-location)


cor_matrix <- cor(data_jun18_wide)

corrplot(cor_matrix, method="color")


data_jun18 %>% count(keyword)
```


```{r}
rquery.cormat(data_jun18_wide)

rcorr_matrix <- rcorr(as.matrix(data_jun18_wide %>% select(!location))) 
M <- rcorr_matrix$r
p_mat <- rcorr_matrix$P
corrplot(M, type = "upper", order = "hclust", p.mat = p_mat, sig.level = 0.01)
corrplot(M)
??rcorr_matrix

signif(rcorr_matrix$r, 2)

chart.Correlation(data_jun18_wide %>% select(!location), histogram=TRUE, pch=19)

cor(as.matrix(data_jun18_wide %>% select(!location)))

data_jun18_wide %>% select(!location)


corr <- cor(data_jun18_wide %>% select(!location), na.rm = TRUE)
ggcorrplot(rcorr_matrix)



corr <- round(cor(data_jun18_wide %>% select(!location)), 1)

rcorr_matrix

chart.Correlation(data_jun18_wide %>% select(!location), histogram=TRUE, pch=19)


row.names(data_jun18_wide) <- data_data_jun18_wide <-  as.data.frame(data_jun18_wide)
jun18_wide$location

data_jun18_wide <- data_jun18_wide %>% select(-location)


data_jun18_wide %>% 
  mutate(hits = replace(hits, is.na(hits), 0)) %>% 


cor(data_jun18_wide, na.rm = TRUE)

?cor
```


```{r}
data_jun18_wide %>% 
  ggplot(aes(`is covid vaccine the mark of the beast`, `biden stole the election`)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

