---
title: "01_Cleaning_Cross_State_Bidaily"
output: html_document
---


```{r}
if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis/"

library(tidyverse)
library(gridExtra)
library(lubridate)
library(readstata13)
library(tidylog)

weekly_cross_dir <- file.path(dropbox_file_path, "Data/google_trends/RawData/cross_state_bidaily/")
state_pop_data <- read.csv(file.path(dropbox_file_path, "Data/city_population/FinalData/brazil_state_pop.csv"))
admin_data <- read.csv(file.path(dropbox_file_path, "Data/brazil_admin_data/brazil_covid19_200419.csv"), encoding = "UTF-8")
brazil_level_data <- 
  read.csv(
    file.path(dropbox_file_path, "Data/google_trends/RawData/brazil_country_extract_2020-05-11.csv"), 
    encoding = "UTF-8"
  )

weekly_df <- 
  list.files(weekly_cross_dir, full.names = TRUE, pattern = "*.csv") %>% 
  map_dfr(read.csv, encoding = "UTF-8")
```

Clean dates var
```{r}
#create var for beginning and ending of the week
weekly_df <- 
  weekly_df %>% 
  separate(dates, into = c("date_beg", "date_end"), sep = " ") %>% 
  mutate_at(vars(date_beg, date_end), as.Date)
```

clean name of the state
```{r}
weekly_df <- 
  weekly_df %>% 
  filter(unit_level == "region") %>% 
  mutate(state = str_remove(location, "State of "))
```

Merging admin data and population data

```{r}
admin_data <- 
  admin_data %>% 
  mutate(
    state_en = stringi::stri_trans_general(state, "Latin-ASCII") %>% str_to_upper(), 
    date = as.Date(date)
  ) 

admin_data <- 
  admin_data %>% 
  left_join(
    state_pop_data, 
    by = c("state" = "State")
  )
```

# Create new variables: death rate, cases & deaths per 100,000

```{r}
admin_data <- 
  admin_data %>% 
  mutate(
    case_rate = cases*100000/estimate_2018_state, 
    death_rate = deaths*100000/estimate_2018_state, 
    fatalities_per_case = deaths/cases
  )

admin_data <- 
  admin_data %>% 
  mutate(fatalities_per_case = if_else(is.na(fatalities_per_case), 0, fatalities_per_case))
```

# Create new variables for growth rate of cases and deaths 

```{r}
admin_data <- 
  admin_data %>%
  group_by(state) %>% 
  mutate(
    diff_date = as.numeric(date - lag(date)), 
    diff_growth_cases = cases - lag(cases), 
    diff_growth_deaths = deaths - lag(deaths), 
    growth_rate_cases = (diff_growth_cases/diff_date)*100/lag(cases),
    growth_rate_deaths = (diff_growth_deaths/diff_date)*100/lag(deaths),
    growth_rate_cases = if_else(is.na(growth_rate_cases), 0, growth_rate_cases), 
    growth_rate_deaths = if_else(is.na(growth_rate_deaths), 0, growth_rate_deaths), 
    growth_rate_cases = if_else(is.infinite(growth_rate_cases), 100, growth_rate_cases), 
    growth_rate_deaths = if_else(is.infinite(growth_rate_deaths), 100, growth_rate_deaths), 
  ) %>% 
  ungroup()
```

#Merging admin data and searches data

Now merging the weekly data of cross state trends with the admin + population data.

We have data on: 
- searches over a 2-day period 
- cases/deaths at a daily level

In order to merge them, we take the first day of the searches, and merge it with the daily admin data. In other words, we keep the date for the 1st of the 2-day period. We complete the date_beg variable so that we can merge it with all the days in the admin data.

```{r}
weekly_df <-
  weekly_df %>%
  mutate(date_beg = as.Date(date_beg)) %>%
  complete(date_beg = seq.Date(min(date_beg), max(date_beg), by="day"))

weekly_admin_df <-
  weekly_df %>%
  full_join(admin_data, by = c("state" = "state", "date_beg" = "date"))

## AN ALTERNATIVE APPROACH 
# we first merge the datasets after the date when the admin data is complete
# weekly_admin_df <- 
#   admin_data %>% 
#   left_join(weekly_df, by = c("state" = "state", "date" = "date_beg")) 
```


# We create categories for key words

For now, we are assigning all other words to 1st person. There is an issue with the encoding that we will have to review for better accuracy of assigning categories. 
```{r}
weekly_admin_df <- 
  weekly_admin_df %>% 
  mutate(
    categories = case_when(
      keyword %in% 
        c( "dificuldade ao respirar", "dor nos olhos", "falta de cheiro", 
           "febre", "febre alta valor","perda de olfato", "tosse",
           "sintomas do coronavirus", "teste de coronavírus", "cansaço") ~ "symptoms",
      keyword %in% 
        c("coronavirus", "covid", "ajuda do coronavírus", "corona") ~ "virus", 
      keyword %in% 
        c("como tratar o coronavírus", "estou com falta de ar",
        "estou com febre", "Eu fico em casa", "eu tenho coronavírus",
        "Perdi o olfato", "quais são os sintomas do coronavírus"
        ) ~ "in_1st_person", 
      TRUE ~ "in_1st_person"
      )
  )
```

We merge the data with the Brazil country-level data, so that we can calculate then the relative value across states and over time

we only need the hits at the national level by date and keyword (we can remove the rest of the columns). 

We do the following: 
- merge the variable hits from the national level with the rest of our data at the state level. Call this hits_national
- then divide the variable 


```{r}
brazil_level_data <- 
  brazil_level_data %>% 
  dplyr::select(date, hits, keyword) %>% 
  rename(hits_national = hits)
  
brazil_level_data <- 
  brazil_level_data %>% 
  mutate(date = as.Date(date))

weekly_admin_df <- 
  weekly_admin_df %>% 
  left_join(
    brazil_level_data, 
    by = c("date_beg" = "date", "keyword")
  )

weekly_admin_df %>% 
  mutate(
    hits_national = if_else(hits_national == "<1", 0.5, hits_national)
  )

    hits_national = if_else(hits_national == 0, 0.1, hits_national)

weekly_admin_df %>% 
  mutate
    hits_national = as.character(hits_national),
    hits_national = if_else(hits_national == "<1", 0.5, hits_national)
  )

    hits_national = if_else(hits_national == 0, 0.1, hits_national) 
  ) %>% 
  count(hits_national)

    hits_national = if_else(hits_national == "<1", 0.5, hits_national))
```


Save data
```{r}
#data of weekly cases
write.csv(weekly_df, file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crossstates_bidaily.csv"))
write_rds(weekly_df, file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crossstates_bidaily.rds"))

#data of weekly cases merged with data of pop + admin cases
write.csv(weekly_admin_df, file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crossstates_bidaily_admin.csv"))
write_rds(weekly_admin_df, file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crossstates_bidaily_admin.rds"))
```

