---
title: "Dashboard Searches"
author: "Manuel Ramos"
date: "04/13/2020"
output: flexdashboard::flex_dashboard
runtime: shiny
resource_files:
- brazil_trends_clean_final.csv
---

```{r include=FALSE}
library(tidyverse)
library(gridExtra)
library(shinydashboard)
library(dashboardthemes)

setwd(dir = "/Users/wb537287/Documents/GitHub/covid-social-media-analysis/DataWork/google_trends/")

#trends_df <- read.csv(file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_trends_clean_final.csv"))

#trends_df <- read.csv("/Users/wb537287/Dropbox/COVID Social Media Analysis/Data/google_trends/FinalData/brazil_trends_clean_final.csv")
#trends_df <- read.csv("/Users/wb537287/Documents/GitHub/covid-social-media-analysis/DataWork/google_trends/brazil_trends_clean_final.csv")
trends_df <- read.csv("brazil_trends_clean_final.csv")


trends_df <- trends_df %>% mutate(date = as.Date(date))
trends_df <- trends_df %>% mutate(state = as.character(state))
```

Searches by State {data-navmenu="Google Trends and Covid-19"}
=====================================

```{r fig.align="center"}
states_name <-
  trends_df %>%
  count(state) %>%
  pull(state)

ui <- dashboardPage(
  dashboardHeader(
    title = h3("Brazil Trends Dashboard")
  ),
  dashboardSidebar(
    selectInput(
      inputId = 'state',
      label = 'State',
      choices = states_name,
      selected = "Acre"
      )
    ),
  dashboardBody(
    plotOutput('plotName'), 
    shinyDashboardThemes(theme = "blue_gradient")
  )
)

server <- function(input, output) {
  dataset <- reactive({
      trends_df %>%
      filter(!is.na(categories), categories != "consequences") %>%
      group_by(categories, state, date) %>%
      mutate(mean_hits = mean(hits, na.rm = TRUE)) %>%
      ungroup()
  })

  output$plotName <- renderPlot({
      dataset() %>%
      filter(state == input$state)  %>%
      ggplot() +
      geom_line(
        aes(date, mean_hits, group = categories, color = fct_reorder2(categories, date, mean_hits)), 
        size = 1.25
      ) +
      geom_line(aes(date, cases, color = "Covid-19 cases"), color = "black", size = 2) +
      coord_cartesian(ylim = c(0, 100)) +
      labs(
        title = "Evolution of Search Trends Compared to Covid-19 Cases",
        color = "Category", 
        caption = "The black line indicates number of cases"
      ) +
      theme_light() +
      theme(
        plot.title = element_text(size = 16, hjust = 0.55), 
        text = element_text(size = 14)
      ) 
  }, 
  height = 700, width = 900
)}

shinyApp(ui = ui, server = server)
```

Presence of Covid-19 per State {data-navmenu="Covid-19 per State"}
=====================================

```{r fig.width=12, fig.height=8}
case_rate_per_state <- 
  trends_df %>% 
  filter(date == "2020-04-01", is.na(keyword) | keyword == "coronavirus") %>% 
  ggplot() +
  geom_col(aes(fct_reorder(state, case_rate), case_rate)) +
  coord_flip() +
  labs(x = "Cases per 100,000 people") +
  theme_light()

death_rate_per_state <- 
  trends_df %>% 
  filter(date == "2020-04-01", is.na(keyword) | keyword == "coronavirus") %>% 
  ggplot() +
  geom_col(aes(fct_reorder(state, death_rate), death_rate)) +
  coord_flip() +
  labs(x = "Deaths per 100,000 people") +
  theme_light()


fatality_per_case_state <- 
  trends_df %>% 
  filter(date == "2020-04-01", is.na(keyword) | keyword == "coronavirus") %>% 
  ggplot() +
  geom_col(aes(fct_reorder(state, fatalities_per_case), fatalities_per_case)) +
  coord_flip() + 
  labs(x = "Deaths per case") +
  theme_light()
  
grid.arrange(case_rate_per_state, death_rate_per_state, fatality_per_case_state, ncol = 3)
```

Overall Cases of Covid-19 per State {data-navmenu="Covid-19 per State"}
=====================================

Column {data-width=600}
-------------------------------------

### Cases per State
```{r}
trends_df %>% 
  filter(!is.na(state), !is.na(cases)) %>% 
  count(cases, date, state) %>% 
  ggplot() +
  geom_line(aes(date, cases, group = state, color = fct_reorder2(state, date, cases))) +
  labs(
    title = "Cases per State",
    color = "State"
  ) +
  theme_light()
```

Column {data-width=600}
-------------------------------------

### Deaths per State
```{r}
trends_df %>% 
  filter(!is.na(state), !is.na(cases)) %>% 
  count(deaths, date, state) %>% 
  ggplot() +
  geom_line(aes(date, deaths, group = state, color = fct_reorder2(state, date, deaths))) +
  labs(
    title = "Deaths per State",
    color = "State"
  ) +
  theme_light()

```

Overall Case Rates of Covid-19 per State {data-navmenu="Covid-19 per State"}
=====================================

Column {data-width=600}
-------------------------------------

### Case Rate per State

```{r}
trends_df %>% 
  filter(!is.na(state), !is.na(case_rate)) %>% 
  count(case_rate, date, state) %>% 
  ggplot() +
  geom_line(aes(date, case_rate, group = state, color = fct_reorder2(state, date, case_rate))) +
  labs(
    title = "Case rate per State",
    y = "Cases per 100,000 people",
    color = "State"
  ) + 
  theme_light()
```

Column {data-width=600}
-------------------------------------

### Death Rate per State

```{r}
trends_df %>% 
  filter(!is.na(state), !is.na(death_rate)) %>% 
  count(death_rate, date, state) %>% 
  ggplot() +
  geom_line(aes(date, death_rate, group = state, color = fct_reorder2(state, date, death_rate))) +
  labs(
    title = "Death rate per State",
    y = "Deaths per 100,000 people",
    color = "State"
  ) + 
  theme_light()
```

Presence of Covid-19 per State {data-navmenu="Overall Google Trends"}
=====================================


```{r, fig.width=12, fig.height=8, fig.align="center"}
trends_df %>% 
  filter(!is.na(state), !is.na(keyword), !is.na(hits)) %>% 
  group_by(date, keyword) %>% 
  summarize(
    weighted_mean_hits = weighted.mean(hits, w = estimate_2018_state)
  ) %>% 
  ggplot() +
  geom_line(aes(date, weighted_mean_hits)) +
  labs(
    y = "Weighted Average Number of Hits",
    title = "Average Number of Hits by Keyword Over Time", 
    subtitle = "The average is weighted by the number of people in each state"
  ) +
  facet_wrap(vars(keyword)) + 
  theme_light()
```
```{r echo=FALSE}
# states_name <- 
#   trends_df %>% 
#   count(state) %>%
#   filter(!state %in% c("Santa Catarina", "Rondônia", NA)) %>% 
#   pull(state)
# 
# ui <- fluidPage(
#   sidebarPanel(
#     selectInput(
#       inputId = 'state', 
#       label = 'State',
#       choices = states_name, 
#       selected = "Acre"
#     ), 
#   mainPanel(
#     plotOutput('plotName')
#   )
#   )
# )
# 
# server <- function(input, output) {
#   dataset <- reactive({
#       trends_df %>% 
#       filter(!is.na(categories), categories != "consequences") %>% 
#       group_by(categories, state, date) %>% 
#       mutate(mean_hits = mean(hits, na.rm = TRUE)) %>% 
#       ungroup()
#   })
#   
#   output$plotName <- renderPlot({ 
#       dataset() %>% 
#       filter(state == input$state)  %>% 
#       ggplot() + 
#       geom_line(aes(date, mean_hits, group = categories, color = categories)) +
#       geom_line(aes(date, cases)) +
#       coord_cartesian(ylim = c(0, 100))
#   }, height = 700, width = 800)
#   
# }
# 
# shinyApp(ui = ui, server = server)
```

