---
title: "06_Time_Lags"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(gridExtra)
library(lubridate)

#trends_df <- read.csv(file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_trends_clean_final.csv"))

trends_df <- read.csv("/Users/wb537287/Dropbox/COVID Social Media Analysis/Data/google_trends/FinalData/brazil_trends_clean_final.csv")
```

```{r}
trends_df <- trends_df %>% mutate(date = as.Date(date))
```

We restrict to keywords with at least 1000 observations
```{r}
trends_df %>% 
  count(keyword)
```
# Overall picture across keyword

All keywords for all states

```{r}
trends_df %>% 
  filter(!is.na(state), !is.na(keyword), !is.na(hits)) %>% 
  group_by(date, keyword, state) %>% 
  summarize(
    weighted_mean_hits = weighted.mean(hits, w = estimate_2018_state)
  ) %>% 
  ggplot() +
  geom_line(aes(date, weighted_mean_hits, group = state, color = state)) +
  labs(
    y = "Weighted Average Number of Hits",
    title = "Average Number of Hits by Keyword Over Time (Symptoms)", 
    subtitle = "The average is weighted by the number of people in each state"
  ) +
  facet_wrap(vars(keyword))
```

# Selecting main words 

Based on the graph above, we select the main keywords 

```{r}
# we select a group of keywords
selected_keywords <- 
  c("ajuda do coronavírus", "cloroquina", "como tratar o coronavírus", 
    "coronavirus", "covid", 
    "Estou com falta de ar", "estou com febre", "febre", 
    "fique em casa", "medicos", "quais são os sintomas do coronavírus", 
    "sintomas do coronavirus", "tosse")

main_words_df <- 
  trends_df %>% 
  filter(keyword %in% selected_keywords)

# we now filter for those keywords which appear in a majority of states

main_keywords <- 
  main_words_df %>% 
  count(keyword) %>% 
  filter(n > 1500) %>% pull(keyword)

main_words_df <- 
  trends_df %>% 
  filter(keyword %in% main_keywords)
```


To begin with, we are going to focus for now on "sintomas do coronavirus" and "tosse", which seem to have sufficient variation yet a clear, increasing trend as well

# We compute the growth rate for cases and deaths of coronavirus

```{r}
trends_df <- 
  trends_df %>% 
  group_by(state, keyword) %>% 
  mutate(
    diff_date = as.numeric(date - lag(date)), 
    diff_growth_cases = cases - lag(cases), 
    diff_growth_deaths = deaths - lag(deaths), 
    diff_growth_hits = hits - lag(hits), 
    growth_rate_cases = (diff_growth_cases/diff_date)*100/lag(cases),
    growth_rate_deaths = (diff_growth_deaths/diff_date)*100/lag(deaths),
    growth_rate_hits = (diff_growth_hits/diff_date)*100/lag(hits), 
    growth_rate_cases = if_else(is.na(growth_rate_cases), 0, growth_rate_cases), 
    growth_rate_deaths = if_else(is.na(growth_rate_deaths), 0, growth_rate_deaths), 
    growth_rate_cases = if_else(is.infinite(growth_rate_cases), 100, growth_rate_cases), 
    growth_rate_deaths = if_else(is.infinite(growth_rate_deaths), 100, growth_rate_deaths), 
    growth_rate_hits = if_else(is.na(growth_rate_hits), 0, growth_rate_hits), 
    growth_rate_hits = if_else(is.infinite(growth_rate_hits), 100, growth_rate_hits)
  ) %>% 
  ungroup()
```

We check how the case rate looks like

```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state)) %>% 
  group_by(categories, state, date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = categories, color = fct_reorder2(categories, date, mean_hits))) +
  geom_line(aes(date, growth_rate_cases)) + 
  labs(color = "Category") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state))
```

We check how the death growth rate looks like

```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state)) %>% 
  group_by(categories, state, date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = categories, color = fct_reorder2(categories, date, mean_hits))) +
  geom_line(aes(date, growth_rate_deaths)) + 
  labs(color = "Category") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state))
```

```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state)) %>% 
  group_by(categories, state, date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = categories, color = fct_reorder2(categories, date, mean_hits))) +
  geom_line(aes(date, growth_rate_hits)) + 
  labs(color = "Category") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state))

trends_df %>% 
  count(keyword, state, date, growth_rate_hits, hits)
```


# Time lags

We now compute the difference between these keywords and the maximum in searches

We first create variables for time periods  
```{r}
trends_df <- 
  trends_df %>% 
  mutate(
    month = month(date), 
    wday = wday(date, label = TRUE), 
    week_number = week(date)
  )
```

We compute the z-score for hits
```{r}
trends_df <- 
  trends_df %>%
   group_by(state, keyword) %>%
   mutate(
     hits_feb = mean(hits[month == 2], na.rm=T),
     hits_feb_sd = sd(hits[month == 2], na.rm=T)
   ) %>%
   ungroup()
    
min_sd <- min(trends_df$hits_feb_sd[trends_df$hits_feb_sd > 0], na.rm=T)

trends_df$hits_feb_sd[trends_df$hits_feb_sd < min_sd] <- min_sd
    
trends_df$zscore <- (trends_df$hits - trends_df$hits_feb) / trends_df$hits_feb_sd
```

Compute weekly average of zscore for hits, then find the week with the maximum z-score
```{r}
trends_df <- 
  trends_df %>% 
  group_by(week_number, state, keyword) %>% 
  mutate(week_zscore = mean(zscore, na.rm = TRUE)) %>% 
  ungroup()

trends_df <- 
  trends_df %>% 
  group_by(state, keyword) %>% 
  mutate(
    max_week_zscore = max(week_zscore), 
    max_week_zscore_bin = if_else(max_week_zscore == week_zscore, 1, 0)
  ) 

# need to review later if this worked well
```

We compute the z-score for cases
```{r}
trends_df <- 
  trends_df %>%
   group_by(state, keyword) %>%
   mutate(
     hits_feb = mean(hits[month == 2], na.rm=T),
     hits_feb_sd = sd(hits[month == 2], na.rm=T)
   ) %>%
   ungroup()
    
min_sd <- min(trends_df$hits_feb_sd[trends_df$hits_feb_sd > 0], na.rm=T)

trends_df$hits_feb_sd[trends_df$hits_feb_sd < min_sd] <- min_sd
    
trends_df$zscore <- (trends_df$hits - trends_df$hits_feb) / trends_df$hits_feb_sd
```

Compute weekly average of zscore for hits, then find the week with the maximum z-score
```{r}
trends_df <- 
  trends_df %>% 
  group_by(week_number, state, keyword) %>% 
  mutate(week_zscore = mean(zscore, na.rm = TRUE)) %>% 
  ungroup()

trends_df <- 
  trends_df %>% 
  group_by(state, keyword) %>% 
  mutate(
    max_week_zscore = max(week_zscore), 
    max_week_zscore_bin = if_else(max_week_zscore == week_zscore, 1, 0)
  ) 

# need to review later if this worked well
```

Once we have the maximum z-score, we calculate the days that pass between when the z-score and the maximum growth rate of cases for the main keyword (sintomas do coronavirus, tosse) per state

We compute the period of maximum growth rate for each state and keyword
```{r}
trends_df <- 
  trends_df %>% 
  group_by(state, keyword) %>% 
  mutate(
    max_growth_cases = max(growth_rate_cases), 
    max_growth_cases_bin = if_else(growth_rate_cases == max_growth_cases, 1, 0)
  )
```

We now compute the difference between the week of maximum z-score and the week of maximum case growth rate

```{r}
trends_df %>% 
  group_by(keyword, state) %>% 
  mutate(diff_zscore_cases = date[max_growth_cases_bin == 1] - date[max_week_zscore_bin == 1])

trends_df <- 
  trends_df %>% 
  group_by(keyword, state) %>% 
  mutate(days_since_max_zscore = date - date[max_week_zscore_bin == 1]) %>% 
  ungroup()

trends_df %>% 
  filter(keyword %in% main_keywords) %>% 
  group_by(state, keyword) %>% 
  summarize(
    date_max_week_zscore = mean(date[max_week_zscore_bin == 1]), 
    date_max_growth_cases = mean(date[max_growth_cases_bin == 1])
  ) 
  
```

We try plotting it

```{r}
trends_df %>% 
  filter(keyword == "tosse", state == "Alagoas") %>% 
  group_by(state, keyword) %>% 
  dplyr::select(date, cases, growth_rate_cases, max_week_zscore)
```

## Main question: how to calculate the maximum period of number of cases? If I use a measure like growth rates, then it will always indicate the first days of cases.
