---
title: "07_Cross_state_Analysis"
author: Manuel Ramos
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(gridExtra)
library(lubridate)
library(ggrepel)

if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis"

states_df <- read.csv(file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crosstates_clean.csv"))
```

We focus on the latest date of cases that we have, April 18. The trends data was extracted in April 20, so this is close enough. 

```{r}
states_df_0418 <- 
  states_df %>% 
  filter(date == "2020-04-18") 

selected_keywords <- 
  c("ajuda do coronavírus", "cloroquina", "como tratar o coronavírus", 
    "coronavirus", "covid", 
    "Estou com falta de ar", "estou com febre", "febre", 
    "fique em casa", "medicos", "quais são os sintomas do coronavírus", 
    "sintomas do coronavirus", "tosse")
```


# Correlation between keywords and case rate

```{r}
states_df_0418 %>%
  filter(!is.na(hits), keyword %in% selected_keywords) %>% 
  ggplot() + 
  geom_point(aes(case_rate, hits)) +
  geom_smooth(aes(case_rate, hits), method = "lm") + 
  facet_wrap(vars(keyword)) + 
  coord_cartesian(ylim = c(0, 100))
```

# Correlation between keywords and death rate

```{r}
states_df_0418 %>%
  filter(!is.na(hits), keyword %in% selected_keywords) %>% 
  ggplot() + 
  geom_point(aes(death_rate, hits)) +
  geom_smooth(aes(death_rate, hits), method = "lm") + 
  facet_wrap(vars(keyword)) +
  coord_cartesian(ylim = c(0, 100)) 

```

# Correlation between average of main keywords and case rate
 
We now explore if the average of tosse, febre and como tratar o coronavirus is a better predictor of cases

```{r}
main_keywords <- c("como tratar o coronavírus", "febre", "tosse")

states_df_0418 %>%
  filter(!is.na(hits), keyword %in% main_keywords) %>% 
  group_by(state) %>%
  summarize(
    average_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, average_hits)) +
  geom_smooth(aes(case_rate, average_hits), method = "lm") + 
  geom_label_repel(
    data = . %>% filter(case_rate > 40 | average_hits > 70), 
    aes(case_rate, average_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) +
  coord_cartesian(ylim = c(0, 100))
```

# Correlation between average of main keywords and death rate

```{r}
main_keywords <- c("como tratar o coronavírus", "febre", "tosse")

states_df_0418 %>%
  filter(!is.na(hits), keyword %in% main_keywords) %>% 
  group_by(state) %>%
  summarize(
    average_hits = mean(hits, na.rm = TRUE), 
    death_rate = mean(death_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(death_rate, average_hits)) +
  geom_smooth(aes(death_rate, average_hits), method = "lm") +
  geom_label_repel(
    data = . %>% filter(death_rate > 40 | average_hits > 70), 
    aes(death_rate, average_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) +  coord_cartesian(ylim = c(0, 100))
```

```{r}
states_df_0418 %>% 
  group_by(keyword) %>% 
  summarise(correlation = cor(case_rate, hits)) %>% 
  filter(!is.na(correlation)) %>% 
  arrange(desc(correlation))
```
 
# Focus on "I can't smell" = "Perdi o olfato" 

```{r}
states_df_0418 %>% 
  filter(!is.na(hits), keyword == "Perdi o olfato") %>% 
  ggplot() + 
  geom_point(aes(case_rate, hits)) +
  geom_smooth(aes(case_rate, hits), method = "lm") + 
  ggrepel::geom_label_repel(aes(case_rate, hits, label = state), hjust=0.5, vjust=0.4) +
  coord_cartesian(ylim = c(0, 100))
```
Maranhao has the top number of hits for "Perdi o olfato", yet we don't see such a high case rate there

```{r}
states_df_0418 %>% 
  filter(!is.na(hits), keyword == "Perdi o olfato") %>% 
  ggplot() + 
  geom_point(aes(death_rate, hits)) +
  geom_smooth(aes(death_rate, hits), method = "lm") + 
  ggrepel::geom_label_repel(aes(death_rate, hits, label = state), hjust=0.5, vjust=0.4) +
  coord_cartesian(ylim = c(0, 100))
```

In fact, the death rate in Maranhao is still quite low. This might indicate that there could be an increase in cases (and potentially deaths) in such state in the near future

