---
title: "08_Maranhao_Searches"
author: Manuel Ramos
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, warning=FALSE, message=FALSE)
```

```{r, include = FALSE}
if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis/"

source("/Users/wb537287/Documents/GitHub/covid-social-media-analysis/_master.R")

library(tidyverse)
library(gridExtra)
library(lubridate)
library(readstata13)

# Data of searches in comparison to RJ
search_df <- 
  read.csv(
    "/Users/wb537287/Dropbox/COVID Social Media Analysis/Data/google_trends/RawData/brazil_extract_extra_words_compareBR-RJ.csv", 
    encoding = "UTF-8")

#time series data for each state and keyword
trends_df <- 
  read.csv("/Users/wb537287/Dropbox/COVID Social Media Analysis/Data/google_trends/FinalData/brazil_trends_clean_final.csv") 

#Cross-sectional data at one point in time 
states_df <- read.csv(file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crosstates_clean.csv"))

#data to merge names and sub_code
data("countries")
states_name <- 
  countries %>% 
  filter(country_code == "BR", sub_code != "") %>% head(27) %>% dplyr::select(-country_code)

#admin data
#admin_data <- read.csv(file.path(dropbox_file_path, "Data/brazil_admin_data/brazil_covid19_200419.csv"), encoding = "UTF-8")
admin_data <- read.csv(file.path(dropbox_file_path, "Data/brazil_admin_data/brazil_covid19_200520.csv"), encoding = "UTF-8")

```

We use two different datasets to evaluate the evolution of searches for the state of Maranhao. 

- The first has searches for all states relative to Rio de Janeiro
- The second has searches for all states relative to their own (within a period of 3 months)

# We start by using the data that compares the states relative to Rio de Janeiro

## Maranhao in comparison to others

```{r}
search_df <-   search_df %>% mutate(date = as.Date(date))
```


```{r}
search_df %>% 
  group_by(geo, keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  geom_point(data = . %>% filter(geo == "BR-MA"), aes(date, mean_hits, group = 1), color = "black", size = 1.5) +
  geom_line(data = . %>% filter(geo == "BR-MA"), aes(date, mean_hits, group = 1), color = "black", size = 1.5) +
  facet_wrap(vars(keyword)) + 
  labs(
    subtitle = "Maranhao is highlighted in black"
  )
```

## Maranhao in comparison to the average of all states

```{r}
search_df %>% 
  group_by(keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1, color = "Average of all")) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = search_df %>% filter(geo == "BR-MA"), aes(date, hits, group = 1, color = "Maranhao")) +
  geom_line(data = search_df %>% filter(geo == "BR-MA"), aes(date, hits, group = 1, color = "Maranhao")) +
  facet_wrap(vars(keyword)) + 
  labs(color = "State")
```


## Maranhao only

```{r}
search_df %>% 
  filter(geo == "BR-MA", !is.na(hits)) %>% 
  group_by(keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1)) + 
  geom_point() + 
  geom_line() +
  facet_wrap(vars(keyword))

```

## Comparing Maranhao to Sao Paulo, Rio de Janeiro 

```{r}
search_df %>% 
  filter(geo %in% c("BR-MA", "BR-RJ", "BR-SP"), !is.na(hits)) %>% 
  group_by(keyword, geo, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  facet_wrap(vars(keyword))

```

# Same analysis but reducing the number of words


```{r}
reduced_keywords <- c("como tratar o coronavírus", "Estou com falta de ar", "febre", "eu tenho coronavírus", "tosse")

search_df %>% 
  filter(keyword %in% reduced_keywords) %>% 
  group_by(geo, keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  geom_point(data = . %>% filter(geo == "BR-MA"), aes(date, mean_hits, group = 1), color = "black", size = 1.5) +
  geom_line(data = . %>% filter(geo == "BR-MA"), aes(date, mean_hits, group = 1), color = "black", size = 1.5) +
  facet_wrap(vars(keyword)) + 
  labs(
    subtitle = "Maranhao is highlighted in black"
  )
```

## Maranhao in comparison to the average of all states

```{r}
reduced_keywords <- c("como tratar o coronavírus", "Estou com falta de ar", "febre", "eu tenho coronavírus", "tosse")

search_df %>% 
  filter(keyword %in% reduced_keywords) %>% 
  group_by(keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1, color = "Average of all")) + 
  geom_point() + 
  geom_line() + 
  geom_point(data = search_df %>% filter(geo == "BR-MA", keyword %in% reduced_keywords), aes(date, hits, group = 1, color = "Maranhao")) +
  geom_line(data = search_df %>% filter(geo == "BR-MA", keyword %in% reduced_keywords), aes(date, hits, group = 1, color = "Maranhao")) +
  facet_wrap(vars(keyword)) + 
  labs(color = "State")
```


## Maranhao only

```{r}
search_df %>% 
  filter(geo %in% c("BR-MA", "BR-RJ", "BR-SP"), !is.na(hits), keyword %in% reduced_keywords) %>% 
  group_by(keyword, geo, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  facet_wrap(vars(keyword))
```

## Comparing Maranhao to Sao Paulo, Rio de Janeiro 

```{r}
search_df %>% 
  filter(geo %in% c("BR-MA", "BR-RJ", "BR-SP"), !is.na(hits)) %>% 
  group_by(keyword, geo, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits, group = 1)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  facet_wrap(vars(keyword))

```

## States showing up "I can't smell"

```{r}
reduced_keywords <- c("como tratar o coronavírus", "Estou com falta de ar", "febre", "eu tenho coronavírus", "tosse", "Perdi o olfato")

reduced_keywords <- c("como tratar o coronavírus",  "eu tenho coronavírus", "Perdi o olfato")

search_df %>% 
  filter(keyword %in% reduced_keywords, geo %in% c("BR-RJ", "BR-MA", "BR-SE")) %>% 
  group_by(geo, keyword, date) %>% 
  summarize(mean_hits = mean(hits, na.rm = TRUE)) %>% 
  ggplot(aes(date, mean_hits)) + 
  geom_point(aes(group = geo, color = geo)) + 
  geom_line(aes(group = geo, color = geo)) + 
  facet_wrap(vars(keyword)) +
  labs(
    x = "Date", 
    y = "Average number of hits per state", 
    color = "State code"
  )
```


# We now use the data for Maranhao relative to its own time trends (instead of relative to Rio de Janeiro)

Cleaning and creation of new variables
```{r}
trends_df <- trends_df %>% mutate(date = as.Date(date))

# We create categories for key words
trends_df <- 
  trends_df %>% 
  mutate(
    categories = case_when(
      keyword %in% 
        c("abuso", "abuso sexual", "desemprego",  "dívida", "Educação online",
          "psicologia", "terapia" ) ~ "consequences",
      keyword %in% 
        c("cama de hospital", "desinfetantes", "hidroxicloroquina", "cloroquina",
          "máscara facial", "termômetros", "ventiladores", "medicos", 
          "profissionais de saúde") ~ "resources",
      keyword %in% 
        c( "dificuldade ao respirar", "dor nos olhos", "falta de cheiro", 
           "febre", "febre alta valor","perda de olfato", "tosse",
           "sintomas do coronavirus", "teste de coronavírus") ~ "symptoms",
      keyword %in% 
        c("distância social","fique em casa", "lavar as mãos", "Isolamento social", 
          "isolamento vertical") ~ "prevention", 
      keyword %in% c("coronavirus", "covid", "ajuda do coronavírus") ~ "virus", 
      keyword %in% 
        c("como tratar o coronavírus", "Estou com falta de ar",
        "estou com febre", "Eu fico em casa", "eu tenho coronavírus",
        "Perdi o olfato", "quais são os sintomas do coronavírus", "volta brasil"
        ) ~ "in_1st_person"
      )
  )
```

## Evolution of key terms for Maranhao  
```{r}
trends_df %>% 
  filter(
    !is.na(state), !is.na(keyword), !is.na(hits), state == "Maranhão", 
    !(categories %in% c("consequences", "prevention"))
  ) %>% 
  group_by(date, keyword, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() +
  geom_line(aes(date, mean_hits, group = 1)) +
  labs(
    y = "Average Number of Hits",
    title = "Average Number of Hits by Keyword Over Time (All)"
  ) +
  facet_wrap(vars(keyword)) 
```

## Evolution of key categories v. cases for Maranhao  

```{r}
trends_df %>% 
  filter(!is.na(categories), categories != "consequences", !is.na(state), state == "Maranhão") %>% 
  group_by(categories, state, date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = categories, color = fct_reorder2(categories, date, mean_hits))) +
  geom_line(aes(date, cases)) +
  labs(color = "Category") +
  coord_cartesian(ylim = c(0, 100)) + 
  labs(subtitle = "The black line represents overall cases")
```
 
# Comparison of Maranhao with other states with a high case rate

We calculate the growth rates
```{r}
trends_df <- 
  trends_df %>% 
  group_by(state, keyword) %>% 
  mutate(
    diff_date = as.numeric(date - lag(date)), 
    diff_growth_cases = cases - lag(cases), 
    diff_growth_deaths = deaths - lag(deaths), 
    diff_growth_hits = hits - lag(hits), 
    growth_rate_cases = (diff_growth_cases/diff_date)*100/lag(cases),
    growth_rate_deaths = (diff_growth_deaths/diff_date)*100/lag(deaths),
    growth_rate_hits = (diff_growth_hits/diff_date)*100/lag(hits), 
    growth_rate_cases = if_else(is.na(growth_rate_cases), 0, growth_rate_cases), 
    growth_rate_deaths = if_else(is.na(growth_rate_deaths), 0, growth_rate_deaths), 
    growth_rate_cases = if_else(is.infinite(growth_rate_cases), 100, growth_rate_cases), 
    growth_rate_deaths = if_else(is.infinite(growth_rate_deaths), 100, growth_rate_deaths), 
    growth_rate_hits = if_else(is.na(growth_rate_hits), 0, growth_rate_hits), 
    growth_rate_hits = if_else(is.infinite(growth_rate_hits), 100, growth_rate_hits)
  ) %>% 
  ungroup()
```

We select key words and 8 states with largest case rate (which includes Maranhao)
```{r}
main_keywords <- 
  c("ajuda do coronavírus", "como tratar o coronavírus", "eu tenho coronavírus",
    "Estou com falta de ar", "estou com febre", "febre", 
    "quais são os sintomas do coronavírus", "Perdi o olfato", 
    "sintomas do coronavirus", "tosse")

top_8_case_rate <- 
  trends_df %>% 
  filter(date == "2020-04-15", is.na(keyword) | keyword == "coronavirus") %>% 
  arrange(desc(case_rate)) %>% 
  head(8) %>% pull(state)
```


```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state), state %in% top_8_case_rate) %>% 
  group_by(keyword, state, week = date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = keyword, color = fct_reorder2(keyword, date, mean_hits))) +
  geom_line(aes(date, growth_rate_cases)) + 
  labs(color = "Keyword") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state)) +
  labs(
    y = "Average hits per category", 
    title = "Growth rate of cases (in black) per state over time in comparison to Google Trends"
  )
```


```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state), state == "Maranhão") %>% 
  group_by(keyword, state, date) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = keyword, color = fct_reorder2(keyword, date, mean_hits))) +
  geom_line(aes(date, growth_rate_cases)) + 
  labs(color = "Keyword") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state)) +
  labs(
    y = "Average hits per category", 
    title = "Growth rate of cases (in black) per state over time in comparison to Google Trends"
  )

```

At the weekly level

```{r}
trends_df %>% 
  filter(!is.na(categories), keyword %in% main_keywords, !is.na(state), state %in% top_8_case_rate) %>% 
  mutate(week_number = week(date)) %>% 
  group_by(keyword, state, week_number) %>% 
  mutate(
    mean_hits = mean(hits, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_line(aes(date, mean_hits, group = keyword, color = fct_reorder2(keyword, date, mean_hits))) +
  geom_line(aes(date, growth_rate_cases)) + 
  labs(color = "Keyword") +
  coord_cartesian(ylim = c(0, 100)) + 
  facet_wrap(vars(state)) +
  labs(
    y = "Average hits per category", 
    title = "Growth rate of cases (in black) per state over time in comparison to Google Trends"
  )

```


# Re-do cross-sectional graphs for mid march, to check if Maranhao was an outlier then

We focus on March 15. We need to extract the hits and the case rate in that date for Maranhao and other states

So we start by merging the case rate with the cases on March 15

```{r}
#we start by merging the search_df with the states name
search_df <- 
  search_df %>% 
  left_join(states_name, by = c("geo" = "sub_code"))

#we now need to merge search_df with the cases over time
search_df <- 
  search_df %>% mutate(date = as.Date(date)) %>% 
  left_join(
    admin_data %>% mutate(date = as.Date(date)), 
    by = c("date" = "date", "name" = "state")
  )

```

