---
title: "09_Cross_state_over_time"
author: Manuel Ramos
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=10, size = 10, background = "light", warning=FALSE, message=FALSE)
```

```{r, include = FALSE}
if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis/"

library(tidyverse)
library(gridExtra)
library(lubridate)
library(readstata13)
library(ggrepel)

sjPlot::set_theme(base = theme_light())


#admin data of cases and deaths
data <- read.csv(file.path(dropbox_file_path, "Data/google_trends/FinalData/brazil_crossstates_weekly_admin.csv"))
```

Create a variable for matching date intermediate of hits and date of covid-19 reporting
```{r}
data <- 
  data %>% 
  mutate(
    date_interm = as.Date(date_interm), 
    date_covid = as.Date(date_covid), 
    date_matched = if_else(date_interm == date_covid, 1L, 0L)
  )
```


# Plotting the figure for the keyword febre
```{r}
data %>% 
  filter(date_matched == 1, date_interm > "2020-02-29", keyword == "febre") %>%  
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 80 | (mean_hits > 60 & date_interm > "2020-04-01")), 
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") + 
  labs(
    caption = "Keywords used: 'febre'"
  )
```

# Doing the same graph, but including now "tosse", "febre", and "como tratar o coronavirus"

```{r}
data %>% 
  filter(date_matched == 1, keyword %in% c("febre", "tosse", "como tratar o coronav<ed>rus"), date_interm > "2020-02-29") %>% 
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 75 | (mean_hits > 60 & date_interm >= "2020-04-01")), 
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") +
  labs(
    caption = "Keywords used: 'febre', 'tosse', and 'como tratar o coronavirus'"
  )
```
 
# Same graph but using all keywords extracted

```{r}
data %>% 
  filter(date_matched == 1, keyword %in% c("febre", "tosse", "como tratar o coronav<ed>rus"), date_interm > "2020-02-29") %>% 
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 75 | (mean_hits > 60 & date_interm >= "2020-04-01")), 
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") +
  labs(
    caption = "Keywords used: 'febre', 'tosse', 'como tratar o coronavirus', 'ajuda do coronavirus', 
    'estou com falta de ar', 'estou com febre', eu tenho coronavirus', 'perdi o olfato', quais son os sintomas do coronavirus'"
  )
```


# Which are the states with the largest case rate on April 18th?

```{r}
data %>% 
  filter(date_covid == "2020-04-18") %>% 
  count(case_rate, state) %>% 
  arrange(desc(case_rate)) %>% 
  ggplot()+ 
  geom_col(aes(fct_reorder(state, case_rate), case_rate)) +
  coord_flip() + 
  labs(
    title = "States ordered by case rate on April 18th, 2020", 
    x = "State", 
    y = "Cases per 100,000 people"
  )
```

# Which are the states with the largest death rate on April 18th?

```{r}
data %>% 
  filter(date_covid == "2020-04-18") %>% 
  count(death_rate, state) %>% 
  arrange(desc(death_rate)) %>% 
  ggplot()+ 
  geom_col(aes(fct_reorder(state, death_rate), death_rate)) +
  coord_flip() + 
  labs(
    title = "States ordered by death rate on April 18th, 2020", 
    x = "State", 
    y = "Deaths per 100,000 people"
  )
```


# Second graph (3 terms) only highlighting the most affected states on April 18th

## Top 6 states
```{r}
top_6_states <- 
  data %>% 
  filter(date_covid == "2020-04-18") %>% 
  count(case_rate, state) %>% 
  arrange(desc(case_rate)) %>% 
  head(6) %>% pull(state) 

data %>% 
  filter(date_matched == 1, keyword %in% c("febre", "tosse", "como tratar o coronav<ed>rus"), date_interm > "2020-02-29") %>% 
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(state %in% top_6_states), 
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") +
  labs(
    caption = "Keywords used: 'febre', 'tosse', and 'como tratar o coronavirus'"
  )

```

## Top 12 states
```{r}
top_12_states <- 
  data %>% 
  filter(date_covid == "2020-04-18") %>% 
  count(case_rate, state) %>% 
  arrange(desc(case_rate)) %>% 
  head(12) %>% pull(state) 

data %>% 
  filter(date_matched == 1, keyword %in% c("febre", "tosse", "como tratar o coronav<ed>rus"), date_interm > "2020-02-29") %>% 
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(state %in% top_12_states), 
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") +
  labs(
    caption = "Keywords used: 'febre', 'tosse', and 'como tratar o coronavirus'"
  )

```

# Plotting the keyword "I can't smell"

```{r}
data %>% 
  filter(date_matched == 1, date_interm > "2020-02-29", keyword == "perdi o olfato") %>%  
  group_by(date_interm, state) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    case_rate = mean(case_rate, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(case_rate, mean_hits)) + 
  geom_smooth(aes(case_rate, mean_hits), method = "lm") + 
  geom_text_repel(
    aes(case_rate, mean_hits, label = state), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(date_interm), scales = "free") + 
  labs(
    caption = "Keywords used: 'perdi o olfato'"
  )
```


# METHODOLOGICAL ASPECTS: 
1. Average of key words + admin cases. 

2. Relative (case rate) v. absolute (number of cases). Graphs seem more consistent with relative impact of the virus. Not with absolute impact. Maybe because google searches look for relative values (e.g. searches out of the total searches?)

3. Selection of keywords - very discretional (in a dashboard it's easy to let the user pick; but in a presentation, it is a very discretionary decision)