---
title: "02_clean_global"
author: Manuel Ramos
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=8, size = 10, background = "light", warning=FALSE, message=FALSE)
```

```{r}
#user's file path
if(Sys.info()[["user"]] == "wb537287") dropbox_file_path <- "/Users/wb537287/Dropbox/COVID Social Media Analysis"

#load libraries
library(tidyverse)
library(gridExtra)
library(lubridate)
library(readstata13)
library(tidylog)
library(ggrepel)

#import datasets
data_por <- 
  read.csv(
    file.path(
      dropbox_file_path,
      "Data_WB_Global/Country_Level_Cleaned/Portuguese_Cleaned/Portuguese_countries_clean.csv"
    ), 
    encoding = "UTF-8"
  )


data_en <- 
  read.csv(
    file.path(
      dropbox_file_path,
      "Data_WB_Global/Country_Level_Cleaned/English_Cleaned/English_countries_clean.csv"
    ), 
    encoding = "UTF-8"
  )

data_sp <- 
  read.csv(
    file.path(
      dropbox_file_path,
      "Data_WB_Global/Country_Level_Cleaned/Spanish_Cleaned/Spanish_countries_clean.csv"
    ), 
    encoding = "UTF-8"
  )
```

Merge datasets
```{r}
data <- rbind(data_en, data_por, data_sp)
```


Clean dates
```{r}
data <- 
  data %>% 
  mutate(date = as.Date(date))

data <- 
  data %>% 
  mutate(
    week_number = week(date), 
    day = day(date), 
    month = month(date), 
    wday = wday(date, label = TRUE)
  )
```

Find countries that show up under "I can't smell"
```{r}
countries_cant_smell <-  
  data %>% 
  filter(keyword %in% c("I can't smell", "Perdi o olfato", "no puedo oler")) %>% 
  filter(!is.na(hits)) %>% 
  count(geo.x) %>%
  pull(geo.x)
```

Find the date in which the country first shows under "I can't smell"
```{r}
countries_cant_smell_date <- 
  data %>% 
  filter(keyword %in% c("I can't smell", "Perdi o olfato", "no puedo oler"), geo.x %in% countries_cant_smell) %>% 
  filter(!is.na(hits), date > "2020-03-01") %>% 
  group_by(geo.x) %>% 
  summarize(
    first_date_cant_smell = min(date)
  )

# Add a ranking
countries_cant_smell_date <- 
  countries_cant_smell_date %>% 
  arrange(first_date_cant_smell) %>% 
  mutate(ranking_cant_smell = rank(first_date_cant_smell))
```

# daily_cases at Country level (En)

```{r}
data %>% 
  ggplot(aes(date, daily_cases, group = geo.x, color = fct_reorder2(geo.x, date, daily_cases))) + 
  geom_line() + 
  labs(color = "Country Code")
```

# daily_deaths at Country level (En)

```{r}
data %>% 
  ggplot(aes(date, daily_deaths, group = geo.x, color = fct_reorder2(geo.x, date, daily_deaths))) + 
  geom_line() + 
  labs(color = "Country Code")
```

# Across countries: in 1st person

```{r}
data %>% 
  filter(categories == "in_1st_person") %>% 
  group_by(geo.x, week_number) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    daily_cases = mean(daily_cases, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(daily_cases, mean_hits, color = geo.x)) + 
  geom_smooth(aes(daily_cases, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 80), 
    aes(daily_cases, mean_hits, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(week_number), scales = "free_x") + 
  labs(
    caption = "Keywords used: 'febre'"
  )
```

# Across countries: in 1st person or symptoms

```{r}
data %>% 
  filter(categories %in% c("in_1st_person", "symptoms")) %>% 
  group_by(geo.x, week_number) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    daily_cases = mean(daily_cases, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(daily_cases, mean_hits, color = geo.x)) + 
  geom_smooth(aes(daily_cases, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 80), 
    aes(daily_cases, mean_hits, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(week_number), scales = "free_x") + 
  labs(
    caption = "Keywords used: 'febre'"
  )
```

* Issue: the US has so many more daily_cases that there is not much variation in the x axis across countries

## Across countries: in 1st person or symptoms (excluding the U.S.)

```{r}
data %>% 
  filter(categories %in% c("in_1st_person", "symptoms"), geo.x != "US") %>% 
  group_by(geo.x, week_number) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    daily_cases = mean(daily_cases, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(daily_cases, mean_hits, color = geo.x)) + 
  geom_smooth(aes(daily_cases, mean_hits), method = "lm") + 
  geom_text_repel(
    data = . %>% filter(mean_hits > 80), 
    aes(daily_cases, mean_hits, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(week_number), scales = "free_x") + 
  labs(
    caption = "Keywords used: 'febre'"
  )
```

# I can't smell at the weekly level

```{r}
data %>% 
  filter(keyword %in% c("I can't smell", "Perdi o olfato", "no puedo oler")) %>% 
  group_by(week_number, geo.x) %>% 
  summarize(
    mean_hits = mean(hits, na.rm = TRUE), 
    daily_cases = mean(daily_cases, na.rm = TRUE)
  ) %>% 
  ggplot() + 
  geom_point(aes(daily_cases, mean_hits, color = geo.x)) + 
  geom_smooth(aes(daily_cases, mean_hits), method = "lm") + 
  geom_text_repel(
    aes(daily_cases, mean_hits, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  facet_wrap(vars(week_number), scales = "free_x") +
  labs(
    caption = "Keywords used: 'I can't smell'"
  )
```

* It's hard for this alarm system to work at the country level, because it's quite likely that the term shows overall at the country level.


# Case rate and death rate per Country

## Case rate: Highlighting the countries that show up under I can't smell at any point
```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(case_rate)) %>% 
  count(case_rate, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, case_rate, group = geo.x, color = fct_reorder2(geo.x, date, case_rate))) +
  geom_line(
    data = . %>% filter(geo.x %in% countries_cant_smell),
    aes(date, case_rate, group = geo.x, color = fct_reorder2(geo.x, date, case_rate)), size = 2) +
  labs(
    title = "Case rate per Country",
    color = "Country"
  )
```

## daily_cases: Highlighting the countries that show up under I can't smell at any point
```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(case_rate)) %>% 
  count(daily_cases, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, daily_cases, group = geo.x, color = fct_reorder2(geo.x, date, daily_cases))) +
  geom_line(
    data = . %>% filter(geo.x %in% countries_cant_smell),
    aes(date, daily_cases, group = geo.x, color = fct_reorder2(geo.x, date, daily_cases)), size = 2) +
  labs(
    title = "daily_cases per Country",
    color = "Countries"
  )
```

## Death rate: Highlighting the countries that show up under I can't smell at any point
```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(death_rate)) %>% 
  count(death_rate, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, death_rate, group = geo.x, color = fct_reorder2(geo.x, date, death_rate))) +
  geom_line(
    data = . %>% filter(geo.x %in% countries_cant_smell),
    aes(date, death_rate, group = geo.x, color = fct_reorder2(geo.x, date, death_rate)), size = 2) +
  labs(
    title = "Death rate per Countries",
    color = "Country"
  )
```

## daily_deaths: Highlighting the countries that show up under I can't smell at any point
```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(daily_deaths)) %>% 
  count(daily_deaths, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, daily_deaths, group = geo.x, color = fct_reorder2(geo.x, date, daily_deaths))) +
  geom_line(
    data = . %>% filter(geo.x %in% countries_cant_smell),
    aes(date, daily_deaths, group = geo.x, color = fct_reorder2(geo.x, date, daily_deaths)), size = 2) +
  labs(
    title = "daily_deaths per Country",
    color = "Country"
  )
```

## Date of showing up "I can't smell"

```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(death_rate), geo.x %in% countries_cant_smell) %>% 
  count(death_rate, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, death_rate, group = geo.x, color = fct_reorder2(geo.x, date, death_rate))) +
  geom_vline(
    data = countries_cant_smell_date, 
    aes(xintercept = first_date_cant_smell, group = geo.x, color = geo.x)
  ) +
  geom_label_repel(
    data = countries_cant_smell_date, 
    aes(first_date_cant_smell, y = 80, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  labs(
    title = "Death rate per Country",
    color = "Country"
  )
```


## Date of showing up "I can't smell" & daily_deaths (instead of death rate)

```{r}
data %>% 
  filter(!is.na(geo.x), !is.na(death_rate), geo.x %in% countries_cant_smell) %>% 
  count(daily_deaths, date, geo.x) %>% 
  ggplot() +
  geom_line(aes(date, daily_deaths, group = geo.x, color = fct_reorder2(geo.x, date, daily_deaths))) +
  geom_vline(
    data = countries_cant_smell_date, 
    aes(xintercept = first_date_cant_smell, group = geo.x)
  ) +
  geom_label_repel(
    data = countries_cant_smell_date, 
    aes(first_date_cant_smell, y = 4500, label = geo.x), 
    hjust=0.5, vjust=0.4
  ) + 
  labs(
    title = "daily_deaths per Country",
    color = "Country"
  )
```



## Graph of current death rate distinguishing by date of I can't smell

```{r}
data %>% 
  left_join(countries_cant_smell_date, by = "geo.x") %>% 
  mutate(cant_smell_appears = if_else(!is.na(first_date_cant_smell), 1L, 0L) %>% as.character()) %>% 
  filter(date == "2020-05-18", !is.na(death_rate)) %>% 
  count(daily_deaths, geo.x, cant_smell_appears, first_date_cant_smell) %>% 
  arrange(desc(daily_deaths)) %>% 
  ggplot()+ 
  geom_col(aes(fct_reorder(geo.x, daily_deaths), daily_deaths, fill = cant_smell_appears)) +
  geom_label_repel(
    aes(
      x = as.character(geo.x), 
      y = daily_deaths, 
      label = as.character(first_date_cant_smell)
    ), 
    position = position_fill(vjust = 0)
  ) + 
  labs(
    color = "Country category"
  ) + 
  coord_flip() + 
  labs(
    title = "Countries ordered by daily_deaths on May 18th, 2020", 
    x = "Country", 
    y = "daily_deaths"
  )
```

```{r}
data %>% 
  left_join(countries_cant_smell_date, by = "geo.x") %>% 
  mutate(cant_smell_appears = if_else(!is.na(first_date_cant_smell), 1L, 0L) %>% as.character()) %>% 
  filter(date == "2020-05-18", !is.na(death_rate)) %>% 
  count(death_rate, geo.x, cant_smell_appears, first_date_cant_smell) %>% 
  arrange(desc(death_rate)) %>% 
  ggplot()+ 
  geom_col(aes(fct_reorder(geo.x, death_rate), death_rate, fill = cant_smell_appears)) +
  geom_label_repel(
    aes(
      x = as.character(geo.x), 
      y = death_rate, 
      label = as.character(first_date_cant_smell)
    ),  
    position = position_fill(vjust = 0.5)
  ) + 
  coord_flip() + 
  labs(
    title = "Countries ordered by death rate on May 18th, 2020", 
    x = "Country", 
    y = "Deaths per 100,000 people"
  )
```


Troubleshooting
```{r}
data_por %>% 
  filter(daily_cases < 0)

```

